Data preparation Study 3 - Merit Scale

### 2021-06-16 paper_study3.Rmd
correlaciones latentes
Se simplifica el código eliminando los puntajes factoriales

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
output:
  bookdown::pdf_document2:
    template: null
    toc: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
#bibliography:
#  - ../input/bib/EconomiaMoral.bib
#  - ../input/bib/meritocracy.bib
#  - ../input/bib/invariance.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

```{r val-conv}
pacman::p_load(kableExtra, polycor)

@@

pl1 <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_1)
pl2 <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_1)
pl3 <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_1)
pl4 <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_1)
valores1 <- round(c(pl1,pl2,pl3,pl4),4)

pl1b <- polycor::polyserial(data_s3_conv$perc_merit,data_s3_conv$get_ah_5)
pl2b <- polycor::polyserial(data_s3_conv$pref_merit,data_s3_conv$get_ah_5)
pl3b <- polycor::polyserial(data_s3_conv$perc_nmerit,data_s3_conv$get_ah_5)
pl4b <- polycor::polyserial(data_s3_conv$pref_nmerit,data_s3_conv$get_ah_5)
valores2 <- round(c(pl1b,pl2b,pl3b,pl4b),4)

pl1c <- cor(data_s3_conv$perc_merit,data_s3_conv$nbmerit)
pl2c <- cor(data_s3_conv$pref_merit,data_s3_conv$nbmerit)
pl3c <- cor(data_s3_conv$perc_nmerit,data_s3_conv$nbmerit)
pl4c <- cor(data_s3_conv$pref_nmerit,data_s3_conv$nbmerit)
valores3 <- round(c(pl1c,pl2c,pl3c,pl4c),4)

variables= c("Meritocratic Perception", "Meritocratic Preferences", "Unmeritocratic Perception","Unmeritocratic Preferences")

table01 <- cbind(variables, valores1, valores2, valores3)

kable(table01, col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"), full_width = T, linesep = "", caption = c("Polyserial and Pearson correlation whit other merit scales")) %>% kable_styling(
    full_width = T,
    position = "center",
    font_size = 14,
    bootstrap_options=c("striped", "bordered")) %>%
  column_spec(column = 1, width = "1.5cm", )

@@

load(file = "input/data/proc/pov01_fact02.RData")

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
output:
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
    template: null
    toc: false
linkcolor: blue
#bibliography:
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

```{r val-conv}
pacman::p_load(kableExtra, lavaan, dplyr,tidyr)

@@

#
model <- '
perc_merit =~  meritv03_perc_effort + meritv03_perc_talent
perc_nmerit=~ meritv03_perc_wpart + meritv03_perc_netw
pref_merit =~ meritv03_pref_effort + meritv03_pref_talent
pref_nmerit=~ meritv03_pref_wpart + meritv03_pref_netw

nbmerit =~  nb_merit_1 + nb_merit_2 + nb_merit_3 + nb_merit_4 + nb_merit_5 + nb_merit_6 + nb_merit_7

perc_merit ~~  get_ah_1
perc_nmerit ~~ get_ah_1
pref_merit ~~ get_ah_1
pref_nmerit ~~ get_ah_1
nbmerit ~~ get_ah_1

perc_merit ~~  get_ah_5
perc_nmerit ~~ get_ah_5
pref_merit ~~ get_ah_5
pref_nmerit ~~ get_ah_5
nbmerit ~~ get_ah_5

perc_merit ~~  nbmerit
perc_nmerit ~~ nbmerit
pref_merit ~~ nbmerit
pref_nmerit ~~ nbmerit

'
fit3_o1=lavaan::cfa(model,data = dat04w3, estimator = "DWLS", ordered = c("meritv03_perc_effort","meritv03_perc_talent",
"meritv03_perc_wpart","meritv03_perc_netw","meritv03_pref_effort","meritv03_pref_talent","meritv03_pref_wpart","meritv03_pref_netw" ,"nb_merit_1" , "nb_merit_2" , "nb_merit_3" , "nb_merit_4" , "nb_merit_5" , "nb_merit_6" , "nb_merit_7","get_ah_1","get_ah_5"))

# extract correlation matrix from lavaan model
correlations = standardizedsolution(fit3_o1) %>% filter(op=="~~") %>% 
                dplyr::select(lhs,rhs,est.std) %>% 
                dplyr::filter(lhs== "perc_merit" | 
                             lhs== "pref_merit" | 
                             lhs== "perc_nmerit"| 
                             lhs== "pref_nmerit") %>%
                pivot_wider(names_from = "rhs", values_from = "est.std") %>% 
                dplyr::select(lhs,get_ah_1, get_ah_5, nbmerit) %>% 
                mutate(merit_scale=
                      case_when(lhs == "perc_merit" ~  "Meritocratic perception",
                                lhs=="perc_nmerit" ~ "Unmeritocratic perception", 
                                lhs =="pref_merit" ~ "Meritocratic preferences" , 
                                lhs =="pref_nmerit" ~ "Unmeritocratic preferences"))%>% 
                select(merit_scale, get_ah_1, get_ah_5,nbmerit)

# Make table 
kable(correlations,col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Wherewithal"), full_width = T, linesep = "", caption = c("Correlation whit other merit scales")) %>% kable_styling(
    full_width = T,
    position = "center",
    font_size = 14,
    bootstrap_options=c("striped", "bordered")) %>%
  column_spec(column = 1, width = "1.5cm", )

@@

load(file = "input/data/proc/data_s3_inv.RData")

### 2021-06-16 processing/preparation_study3.rmd
correlaciones latentes
Se simplifica el código eliminando los puntajes factoriales

## Libraries
```{r}
pacman::p_load(haven, sjPlot, dplyr, mice, lavaan, stargazer, corrplot, polycor, knitr, kableExtra, summarytools, sjlabelled)
```

@@

```{r}
dat04w3 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola3.sav",verbose = F)
dim(data) 
```

## Create Perceptions and Preferences for Meritocracy (PPM-S) factor scores

@@

```{r results='hide'}
dat04w3

dat04w3$meritv03_perc_effort <- as.numeric(dat04w3$meritv03_perc_effort)
dat04w3$meritv03_perc_talent <- as.numeric(dat04w3$meritv03_perc_talent)
dat04w3$meritv03_perc_netw <- as.numeric(dat04w3$meritv03_perc_netw)
dat04w3$meritv03_perc_wpart <- as.numeric(dat04w3$meritv03_perc_wpart)
dat04w3$meritv03_pref_effort <- as.numeric(dat04w3$meritv03_pref_effort)
dat04w3$meritv03_pref_talent <- as.numeric(dat04w3$meritv03_pref_talent)
dat04w3$meritv03_pref_wpart <- as.numeric(dat04w3$meritv03_pref_wpart)
dat04w3$meritv03_pref_netw <- as.numeric(dat04w3$meritv03_pref_netw)

model04 <- '
perc_merit =~  meritv03_perc_effort + meritv03_perc_talent
perc_nmerit=~ meritv03_perc_wpart + meritv03_perc_netw
pref_merit =~ meritv03_pref_effort + meritv03_pref_talent
pref_nmerit=~ meritv03_pref_wpart + meritv03_pref_netw'

# generate 5 multiple complete datasets
out <- mice(dat04w3[23:30], m = 5)

D1 <- complete(out, 1)
D2 <- complete(out, 2)
D3 <- complete(out, 3)
D4 <- complete(out, 4)
D5 <- complete(out, 5)

# fit model for each complete dataset
fit1 <- cfa(model04, data = D1)
fit2 <- cfa(model04, data = D2)
fit3 <- cfa(model04, data = D3)
fit4 <- cfa(model04, data = D4)
fit5 <- cfa(model04, data = D5)

# predict scores for all models
p1 <- predict(fit1)
p2 <- predict(fit2)
p3 <- predict(fit3)
p4 <- predict(fit4)
p5 <- predict(fit5)

# compute 'average' across 5 sets of scores:
scores <- (p1 + p2 + p3 + p4 + p5) / 5

#Generate factor scores
scores_dat = as.data.frame(scores)

# Adaptar base para agregar factor scores (sacar todos NA)
dat04w3 = dat04w3[rowSums(is.na(dat04w3[23:28])) != 5, ]

# Merge with factor scores
dat04w3 = cbind(dat04w3, scores)

# Check
stargazer(dat04w3[, 78:81], type = "text")

@@

## Create The Neoliberal Beliefs Inventory (NBI) factor scores

The following code creates new variables corresponding to the four dimensions of the NBI scale. Based on the measurement model, four factorial scores are created. To retrieve cases the "mice" function is used
```{r results='hide'}
datnbmerit <- dat04w3 %>% select(nb_merit_1,
nb_merit_2,
nb_merit_3,
nb_merit_4,
nb_merit_5,
nb_merit_6,
nb_merit_7,
nb_merit_8)
psych::alpha(datnbmerit)

dat04w3$nb_merit_1 <- as.numeric(dat04w3$nb_merit_1)
dat04w3$nb_merit_2 <- as.numeric(dat04w3$nb_merit_2)
dat04w3$nb_merit_3 <- as.numeric(dat04w3$nb_merit_3)
dat04w3$nb_merit_4 <- as.numeric(dat04w3$nb_merit_4)
dat04w3$nb_merit_5 <- as.numeric(dat04w3$nb_merit_5)
dat04w3$nb_merit_6 <- as.numeric(dat04w3$nb_merit_6)
dat04w3$nb_merit_7 <- as.numeric(dat04w3$nb_merit_7)
dat04w3$nb_merit_8 <- as.numeric(dat04w3$nb_merit_8)

model04 <- '
nbmerit =~  nb_merit_1+
nb_merit_2+
nb_merit_3+
nb_merit_4+
nb_merit_5+
nb_merit_6+
nb_merit_7'

# generate 5 multiple complete datasets
out <- mice(dat04w3[57:64], m = 5)

D1 <- complete(out, 1)
D2 <- complete(out, 2)
D3 <- complete(out, 3)
D4 <- complete(out, 4)
D5 <- complete(out, 5)

# fit model for each complete dataset
fit1 <- cfa(model04, data = D1)
fit2 <- cfa(model04, data = D2)
fit3 <- cfa(model04, data = D3)
fit4 <- cfa(model04, data = D4)
fit5 <- cfa(model04, data = D5)

# predict scores for all models
p1 <- predict(fit1)
p2 <- predict(fit2)
p3 <- predict(fit3)
p4 <- predict(fit4)
p5 <- predict(fit5)

# compute 'average' across 5 sets of scores:
scores <- (p1 + p2 + p3 + p4 + p5) / 5

#Generate factor scores
scores_dat = as.data.frame(scores)

# Adaptar base para agregar factor scores (sacar todos NA)
dat04w3 = dat04w3[rowSums(is.na(dat04w3[57:64])) != 5, ]

# Merge with factor scores
dat04w3 = cbind(dat04w3, scores)

data_s3_conv <- dat04w3 %>% select(nbmerit, perc_merit,pref_merit,perc_nmerit,pref_nmerit,get_ah_1, get_ah_5) %>% na.omit()

## Labels

### Variable labels

Created variables are labeled and described

@@

set_label(data_s3_conv$nbmerit)  <- "Personal wherewithal - NBI"
set_label(data_s3_conv$perc_merit)  <- "Meritocratic perception"
set_label(data_s3_conv$pref_merit)   <- "Meritocratic preferences"
set_label(data_s3_conv$perc_wpart)    <- "Unmeritocratic perception"
set_label(data_s3_conv$pref_nmerit)  <-  "Unmeritocratic preferences"

set_label(data_s3_conv$get_ah_1)  <- "Importance of a rich family to get ahead"
set_label(data_s3_conv$get_ah_5)   <- "Importance of hard work to get ahead"

sjmisc::descr(data_s3_conv)

## Final summary
```{r}
sjmisc::descr(data_s3_conv,
      show = c("label","range", "mean", "sd", "NA.prc", "n"))%>%
      kable(.,"markdown")

print(dfSummary(data_s3_conv, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")

@@

save(data_s3_conv,file = "input/data/proc/data_s3_conv.RData")
```

>>>>

## Libraries
```{r}
pacman::p_load(haven, sjPlot, dplyr, tidyr, mice, lavaan, stargazer, corrplot, polycor, knitr, kableExtra, summarytools, sjlabelled)
```

@@

```{r}
dat04w3 <- sjlabelled::read_spss("input/data/original/Estudio_3_ola3.sav",verbose = F)
dim(dat04w3) 
```

## Estimate the measurement model with the dimensions of the merit-scale and other meritocracy scales

@@

```{r results='hide'}

model <- '
perc_merit =~  meritv03_perc_effort + meritv03_perc_talent
perc_nmerit=~ meritv03_perc_wpart + meritv03_perc_netw
pref_merit =~ meritv03_pref_effort + meritv03_pref_talent
pref_nmerit=~ meritv03_pref_wpart + meritv03_pref_netw

nbmerit =~  nb_merit_1 + nb_merit_2 + nb_merit_3 + nb_merit_4 + nb_merit_5 + nb_merit_6 + nb_merit_7

perc_merit ~~  get_ah_1
perc_nmerit ~~ get_ah_1
pref_merit ~~ get_ah_1
pref_nmerit ~~ get_ah_1
nbmerit ~~ get_ah_1

perc_merit ~~  get_ah_5
perc_nmerit ~~ get_ah_5
pref_merit ~~ get_ah_5
pref_nmerit ~~ get_ah_5
nbmerit ~~ get_ah_5

perc_merit ~~  nbmerit
perc_nmerit ~~ nbmerit
pref_merit ~~ nbmerit
pref_nmerit ~~ nbmerit

'

fit3_c1=lavaan::cfa(model,data = dat04w3)

fit3_o1=lavaan::cfa(model,data = dat04w3, estimator = "DWLS", ordered = c("meritv03_perc_effort","meritv03_perc_talent",
"meritv03_perc_wpart","meritv03_perc_netw","meritv03_pref_effort","meritv03_pref_talent","meritv03_pref_wpart","meritv03_pref_netw" ,"nb_merit_1" , "nb_merit_2" , "nb_merit_3" , "nb_merit_4" , "nb_merit_5" , "nb_merit_6" , "nb_merit_7","get_ah_1","get_ah_5"))

cargas1 = left_join(x = standardizedsolution(fit3_c1) %>% filter(op=="=~") %>% dplyr::select(lhs,rhs,est.std),y = standardizedsolution(fit3_o1) %>% filter(op=="=~") %>% dplyr::select(lhs,rhs,est.std),c("lhs","rhs"))

kable(cargas1,
      format = "html",digits = 2, booktabs=T, caption = "Cargas factoriales",escape = FALSE) %>%  
   kable_styling(full_width = T, font_size = 8,  bootstrap_options=c("striped", "bordered")) %>%
     column_spec(column = 1,width = "2 cm") %>% 
   column_spec(column = 2,width = "11 cm") %>%
  column_spec(column = 3,width = "1 cm") %>%
  column_spec(column = 4,width = "1 cm") %>%
      collapse_rows(columns = 1,valign = "middle")

pacman::p_load(dplyr, kableExtra, knitr)

# extract fit indices from models and add to table
sum_fit1<- dplyr::bind_rows(fitmeasures(fit3_c1)[c("chisq","df","cfi","tli","rmsea","srmr")],
                    fitmeasures(fit3_o1)[c("chisq","df","cfi","tli","rmsea","srmr")])

# Customize object
sum_fit1$mod <- c("Continuous","Ordinals")
sum_fit1$nobs <- c(nobs(fit3_c1),nobs(fit3_o1))
sum_fit1$est <- c("MLR","DWLS")
sum_fit1 <- dplyr::select(sum_fit1,mod,nobs,est,everything())
colnames <- c("Treated as","$N$","Estimator","$\\chi^2$","df","CFI","TLI","RMSEA","SRMR")

# Create table
table_cfa_fits1 <-kable(sum_fit1, format="html", digits=3, booktabs=T, col.names=colnames, escape = FALSE) %>% kable_styling(full_width = F, font_size = 10)
table_cfa_fits1

@@

## Calculate the correlations between the scale of this work and other measures of meritocratic beliefs.

@@

correlaciones = standardizedsolution(fit3_o1) %>% filter(op=="~~") %>% 
                dplyr::select(lhs,rhs,est.std) %>% 
                dplyr::filter(lhs== "perc_merit" | 
                             lhs== "pref_merit" | 
                             lhs== "perc_nmerit"| 
                             lhs== "pref_nmerit") %>%
                pivot_wider(names_from = "rhs", values_from = "est.std") %>% 
                dplyr::select(lhs,get_ah_1, get_ah_5, nbmerit) %>% 
                mutate(merit_scale=
                      case_when(lhs == "perc_merit" ~  "Meritocratic perception",
                                lhs=="perc_nmerit" ~ "Unmeritocratic perception", 
                                lhs =="pref_merit" ~ "Meritocratic preferences" , 
                                lhs =="pref_nmerit" ~ "Unmeritocratic preferences"))%>% 
                select(merit_scale, get_ah_1, get_ah_5,nbmerit)

kable(correlaciones,col.names = c("Merit-Scale","Social Origin ", "Hard Work " , "Personal wherewithal - NBI"),
      format = "html",digits = 2, booktabs=T,escape = FALSE) %>%  
      kable_styling(full_width = T, font_size = 8,  bootstrap_options=c("striped", "bordered")) %>%
      column_spec(column = 1,width = "2 cm") %>% 
      column_spec(column = 2,width = "11 cm") %>%
      column_spec(column = 3,width = "1 cm") %>%
      collapse_rows(columns = 1,valign = "middle")
```

## Final summary
```{r}
#sjmisc::descr(dat04w3,
#      show = c("label","range", "mean", "sd", "NA.prc", "n"))%>%
#      kable(.,"markdown")

print(dfSummary(dat04w3, valid.col = FALSE, graph.magnif = 0.75),
      max.tbl.height = 300, method = "render")

@@

save(dat04w3,file = "input/data/proc/data_s3_conv.RData")
```
